Effects of epidemiological structure on the transient evolution of HIV virulence

### 2016-05-06 ms/ms.Rnw
more bib updates

\todo{I think we need a section here to describe partnership dynamics/contact structure, i.e. define the $\rho$, $c_u$, $c_e$, $c_w$ parameters - will be similar to \champr}
\input{ms_parms.tex}

@@

Table~\ref{table:parmsTable} gives the ranges of the model parameters used for LHS. Parameter ranges regarding contact and partnership dynamics ($\rho$, $c$, and $c_e/c_w$) are taken from \champr, whereas those regarding infection ($\beta_P$, $D_P$, $\beta_D$, and $D_D$) are taken from \cite{hollingsworth_hiv1_2008}. The remaining parameters are taken from \shir \todo{check: redundant with description a couple of paras above?}

@@

\todo{add Shirreff's parameters to table 1: (\sout{Beta1 = 2.76, Duration1 = 0.25}, Beta3 = 0.76, Duration3 = 0.75, \sout{cmeanbase = 1.25, alphaDist = c(min = 2, max = 7,  delta = 0.1)}).}

@@

Our chosen summary statistics (peak time, peak virulence, equilibrium
virulence, and relative peak virulence) all vary considerably across models
\ref{fig:unidist}.
We first consider the models of intermediate realism (implicit,
instantaneous-switching with and without extra-pair contact, and
pair formation without extra-pair contact). Some parameter
sets for these models lead to low equilibrium virulence ($\approx 2.5 \log_{10} SPVL$)
these same sets lead to correspondingly low
peak virulence ($< 3.5 \log_{10} SPVL$) and early peak times (before 200 years), 
but high relative peaks ($>1.3$)
(Figure~\ref{fig:pairplot}, leftmost column) because the equilibrium virulence is low.
At the opposite extreme, parameter sets that produce high equilibrium virulence 
also produce late peaks ($> 200 \text{years}$), 
high peak virulence, and low relative peaks ($\approx 1.05$).
The pair-formation without extra-pair contact and implicit models
occasionally have parameter sets that select for such low virulence across
the board that they never exceed their initial virulence, leading to a tail
of peak times near 0.

@@

@@

<<plot_sens,dev="png",echo=FALSE,fig.width=10,fig.height=6>>=
ggplot(mL2,aes(LHSval,sumval,colour=model))+
    geom_point(pch=".",alpha=0.5)+
    facet_grid(sumvar~LHSvar,scales="free")+
    geom_smooth(se=FALSE)+labs(x="",y="")+
    scale_x_log10()
@

Figure to-do: maybe go back to bw theme?
\begin{itemize}
\item fig 1. tweak size? remove redundant y axes/labels?
\item fig 2. 6-panel figure? log scale? boundary lines instead of ribbons for edges of envelopes? move labels away from axes?
\item fig 3. change order (Random after pairform)
\item fig 4. add legend, fix axis labels; maybe transparent
\item fig 5. sensitivity plot (unscaled plot); take out plots that are not related ... implicit model doesn't use rho, cu or ce; random doesn't use rho, cmean?, cu, ce; instswitch doesn't have rho or cu or ce; models without epc don't have ce or cu; instswitch + epc doesn't have rho and cu (no uncoupled compartment). Maybe squash panels together, higher resolution

>>>>

\input{ms_parms.tex}

@@

Table~\ref{table:parmsTable} gives the ranges of the model parameters used for LHS. Parameter ranges regarding contact and partnership dynamics ($\rho$, $c$, and $c_e/c_w$) are taken from \champr, whereas those regarding infection ($\beta_P$, $D_P$, $\beta_D$, and $D_D$) are taken from \cite{hollingsworth_hiv1_2008}. The remaining parameters are taken from \shir.

@@

@@

Our chosen summary statistics (peak time, peak virulence, equilibrium
virulence, and relative peak virulence) all vary considerably across models
\ref{fig:unidist}.
We first consider the models of intermediate realism (implicit,
instantaneous-switching with and without extra-pair contact, and
pair formation without extra-pair contact). Some parameter
sets for these models lead to low equilibrium virulence ($\approx 2.5 \log_{10}$ SPVL)
these same sets lead to correspondingly low
peak virulence ($< 3.5 \log_{10}$ SPVL) and early peak times (before 200 years), 
but high relative peaks ($>1.3$)
(Figure~\ref{fig:pairplot}, leftmost column) because the equilibrium virulence is low.
At the opposite extreme, parameter sets that produce high equilibrium virulence 
also produce late peaks ($> 200 \text{ years}$), 
high peak virulence, and low relative peaks ($\approx 1.05$).
The pair-formation without extra-pair contact and implicit models
occasionally have parameter sets that select for such low virulence across
the board that they never exceed their initial virulence, leading to a tail
of peak times near 0.

@@

\emph{why implicit = pair formation? when we increased the unadjusted partnership
mixing rates, the two models started to look different; so they may just be
at some lower bound ...}

@@

<<plot_sens,dev="png",echo=FALSE,fig.width=10,fig.height=6>>=
subset(mL2,
       !(model=="implicit" &
           LHSvar %in% c("rho_base","c_u_ratio","c_e_ratio")))
ggplot(mL2,aes(LHSval,sumval,colour=model))+
    geom_point(pch=".",alpha=0.5)+
    facet_grid(sumvar~LHSvar,scales="free")+
    geom_smooth(se=FALSE)+labs(x="",y="")+
    scale_x_log10()
@

Figure to-do: maybe go back to bw theme?
\begin{itemize}
\item fig 1. tweak size? remove redundant y axes/labels?
\item fig 2. 6-panel figure? log scale? boundary lines instead of ribbons for edges of envelopes? move labels away from axes?
\item fig 3. change order (flip instswitch and instswitch + epd)
\item fig 4. add legend, maybe transparent
\item fig 5. sensitivity plot (unscaled plot); take out plots that are not related ... implicit model doesn't use rho, cu or ce; random doesn't use rho, cmean?, cu, ce; instswitch doesn't have rho or cu or ce; models without epc don't have ce or cu; instswitch + epc doesn't have rho and cu (no uncoupled compartment). Maybe squash panels together, higher resolution

### 2016-05-06 ms/ms_parms.tex
more bib updates

\begin{table}[ht]
\caption{Parameter ranges/values.  Note that $c$ and $\rho$ values are doubled from \champr\  because we keep track of individuals, while they keep track of couples. Starred (*) parameters (used in Figure~\ref{fig:panel3}) are from \shir.}
\centering
\begin{tabular}{c p{2in} c l}
\hline 
Notation & Description & Range/Value & Source\\
\hline % inserts single horizontal line
$\rho$ & Partnership formation rate & 1/10-2/5 & \cite{champredon_hiv_2013} \\
$c$ & Partnership dissolution rate & 1/15-1/5 (1.25*) & \cite{champredon_hiv_2013} \\
$c_u/c_w$ & Relative contact rate for uncoupled transmission & 1/5-5 & Assumption \\
$c_e/c_w$ & Relative contact rate extra-couple & 0.01-1 & \cite{champredon_hiv_2013} \\
$\beta_P$ & Rate of transmission during primary infection & 1.31-5.09 (2.76*) & \cite{hollingsworth_hiv1_2008} \\
$\beta_D$ & Rate of transmission during high transmission disease stage & 0.413-1.28 & \cite{hollingsworth_hiv1_2008} \\
$D_P$ & Duration of primary infection & 1.23/12-6/12 (0.25*) & \cite{hollingsworth_hiv1_2008,shirreff_transmission_2011} \\
$D_D$ & Duration of high transmission disease stage & 4.81/12-14/12 & \cite{hollingsworth_hiv1_2008} \\
$\tsub{\beta}{max}$ & Maximum rate of transmission during asymptomatic stage & 0.317 & \cite{shirreff_transmission_2011} \\
$\beta_{50}$ & SPVL at which infectiousness is half maximum & 13938 & \cite{shirreff_transmission_2011} \\
$\beta_k$ & Hill coefficient: steepness of increase in infectiousness as a function of SPVL & 1.02 & \cite{shirreff_transmission_2011} \\
$\tsub{D}{max}$ & Duration of primary infection & 25.4 & \cite{shirreff_transmission_2011} \\
$D_{50}$ & SPVL at which duration of asymptomatic infection is half maximum & 3058 & \cite{shirreff_transmission_2011} \\
$D_{k}$ & Hill coefficient: steepness of decrease in duration as a function of SPVL & 0.41 & \cite{shirreff_transmission_2011} \\
$\sigma_M$ & Mutation standard deviation of $\log_{10}$ SPVL & 0.12 & \cite{shirreff_transmission_2011} \\
$\tsub{\alpha}{min}$ & Minimum $\log_{10}$ SPVL & 2 & \cite{shirreff_transmission_2011}\\
$\tsub{\alpha}{max}$ & Maximum $\log_{10}$ SPVL & 7 & \cite{shirreff_transmission_2011}\\
$n$ & Number of strains & 21 (51*) & Assumption\\
\hline
\end{tabular}
\label{table:parmsTable}
\end{table}

>>>>

\begin{table}[ht]
\caption{Parameter ranges/values.  Note that $c$ and $\rho$ values are doubled from \champr\  because we keep track of individuals, while they keep track of couples. Starred (*) parameters (used in Figure~\ref{fig:panel3}) are from \shir.}
\centering
\begin{tabular}{c p{2in} c l}
\hline 
Notation & Description & Range/Value & Source\\
\hline % inserts single horizontal line
$\rho$ & Partnership formation rate & 1/10-2/5 & \cite{champredon_hiv_2013} \\
$c$ & Partnership dissolution rate & 1/15-1/5 (1.25*) & \cite{champredon_hiv_2013} \\
$c_u/c_w$ & Relative contact rate for uncoupled transmission & 1/5-5 & Assumption \\
$c_e/c_w$ & Relative contact rate extra-couple & 0.01-1 & \cite{champredon_hiv_2013} \\
$\beta_P$ & Rate of transmission during primary infection & 1.31-5.09 (2.76*) & \cite{hollingsworth_hiv1_2008} \\
$\beta_D$ & Rate of transmission during high transmission disease stage & 0.413-1.28 (0.76*) & \cite{hollingsworth_hiv1_2008} \\
$D_P$ & Duration of primary infection & 1.23/12-6/12 (0.25*) & \cite{hollingsworth_hiv1_2008,shirreff_transmission_2011} \\
$D_D$ & Duration of high transmission disease stage & 4.81/12-14/12 (0.75*) & \cite{hollingsworth_hiv1_2008} \\
$\tsub{\beta}{max}$ & Maximum rate of transmission during asymptomatic stage & 0.317 & \cite{shirreff_transmission_2011} \\
$\beta_{50}$ & SPVL at which infectiousness is half maximum & 13938 & \cite{shirreff_transmission_2011} \\
$\beta_k$ & Hill coefficient: steepness of increase in infectiousness as a function of SPVL & 1.02 & \cite{shirreff_transmission_2011} \\
$\tsub{D}{max}$ & Duration of primary infection & 25.4 & \cite{shirreff_transmission_2011} \\
$D_{50}$ & SPVL at which duration of asymptomatic infection is half maximum & 3058 & \cite{shirreff_transmission_2011} \\
$D_{k}$ & Hill coefficient: steepness of decrease in duration as a function of SPVL & 0.41 & \cite{shirreff_transmission_2011} \\
$\sigma_M$ & Mutation standard deviation of $\log_{10}$ SPVL & 0.12 & \cite{shirreff_transmission_2011} \\
$\tsub{\alpha}{min}$ & Minimum $\log_{10}$ SPVL & 2 & \cite{shirreff_transmission_2011}\\
$\tsub{\alpha}{max}$ & Maximum $\log_{10}$ SPVL & 7 & \cite{shirreff_transmission_2011}\\
$n$ & Number of strains & 21 (51*) & Assumption\\
\hline
\end{tabular}
\label{table:parmsTable}
\end{table}
